---
title: "Aides Drug"
author: "Joel Calderin"
date: "Fall 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
library(dplyr)
druggies = filter(triUmp, triUmp$DrugUser == 'yes')
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$ There is no difference between the new preventative aids drug treatment and the effect it has on drug users

$H_A$ There is a difference between the new preventative aids drug treatment and how it effects drug users


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

Both our response variable and explanatory variable are categorical. Therefore we will use the cat~cat helper file. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=druggies, type="percent")
```


The bar graph displays the percentage of people with aids that got the placebo drug is significantly higher than the people with aids that received the preventative drug. This means you are less likely to get aids with the preventative drug.  


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~infected + treatment, data=druggies)
rowPerc(table1)
```

The table shows that the placebo group was roughly 3 times more likely to get infected compared to the drug group. About 75% of the infected group was given the placebo shot, while only about 25 percent of the infected group was given the vaccine. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The P-Value is 0.0000001242 which is well below 0.05. This means we reject the NULL Hypothesis. 


```{r}
druggies$infected = factor(druggies$infected, c('none', 'aids'))
fisher.test(druggies$treatment,druggies$infected)
```

The P value in this test is equal to .0000000423. The people who got the vaccine are almost 3 times more protected from aids compared to the placebo vaccinated people.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


Is the vaccine effective in preventing aids with the new preventative drug? Yes, the vaccine is effective in preventing aids. Our graphical data showed the significant difference in the people with aids with the vaccine compared to the placebo vaccine. Our numerical and inferential data also tells us that the vaccine was 3 times more protective to aids than the placebo vaccine. We conducted a Chisq test which gave us a P-Value less than 0.05; which means we rejected the NULL Hypothesis. In conclusion, the new preventative drug treatment is effective against the infection aids. 

